---
title: 'Module 1 Project 1: Data Analysis For Sales And Temperature Data, Student Dataset'
author: "Abhilash Kumar Dikshit"
date: "2022-09-28"
bibliography: references.bib
biblio-style: "apalike"
link-citations: true
output:
  pdf_document: default
  html_document:
    df_print: paged
---
# Install VCD Package
```{r}
#install.packages('vcd')
```
# Import VCD Library
```{r}
library(vcd)
```
# Plot: Sales And Temperature Data
```{r}
Sales = c(8,11,15,20,21,11,18,10,6,22)
Temperature = c(69,80,77,84,80,77,87,70,65,90)

#Checking Length
length(Sales)
length(Temperature)
plot(Sales,Temperature,main="Sales And Temperature Data",sub="SCATTER PLOT REPRESENTATION",
     cex.main=2,xlab="Sales Data",ylab="Temperature Data",col="red",cex=2)
```
# Mean Temperature 
```{r}
mean(Temperature)
```

# Delete 3rd element From Sales Vector
```{r}
Sales_Del = Sales[-3]
Sales_Del
```
# Insert Number 16 as 3rd element Into Sales Vector
```{r}
Sales_replace = replace(Sales,3,16)
Sales_replace
```
# Create Vector <names> With Elements "Tom, Dick, Harry"
```{r}
names = c("Tom","Dick","Harry")
names
```
# Create Matrix Of 10 integers (5 row, 2 column)
```{r}
create_matrix = matrix(c(1:10), nrow = 5, ncol = 2)
create_matrix
```
# Create Data Frame <icSales> With Sales And Temperature Attributes
```{r}
icSales = data.frame (Sales,Temperature)
icSales
icSales_New = str(icSales)
```
# Summary Of icScales Data Frame
```{r}
summary_output = summary(icSales)
summary_output
```
# Import Dataset "Student.csv"
```{r}
Student_Data = read.csv("Student.csv")
Student_Data
```
# Display Variable Names Of "Student.csv" Dataset
```{r}
display_var = ls(Student_Data)
display_var
```
# Summary Of Student_Data Dataset
```{r}
summary_student = summary(Student_Data)
summary_student
```
# Summary of the information I learned about the data sets based on the given instructions
Working with R on *MacOS* was initially a bit tough but after completing this assignment, I feel the hands are nicely set on the keyboard and flows in the same direction as my *Goal* towards being a good *Data Scientist*. 

However, there were few hurdles which I faced during the process, and, after looking it up on *web* and spending a good sleepless night, I did fixed many issues and errors w.r.t the packages which was required to complete the given assignment. I have added the *References* section which I find really useful to give a gist of the R concept and how to deal with the real world data. Moreover, I learned *keyboard shortcuts* which is really necessary for R to speed up the day-to-day tasks.

# Install and Import "yihui_pkgs" & "TinyTax" Package
```{r}
#tinytex::install_tinytex()
#tinytex:::install_yihui_pkgs()
library(tinytex)
#tinytex::pdflatex('Dikshit_M1_Project1.tex')
```

# References
<div id="refs">@R-Markdown;@R-Graphics;@R-Stack;@R-Tutorials</div>
# Appendix

```{r code=readLines(knitr::purl('~/Dikshit_M1_Project1.Rmd', documentation = 0)), eval = FALSE}
```


