---
title: "ALY6070:Airbnb New York Analysis (2011 - 2019)"
subtitle: "Week 3"
author: "Group 7"
date: "2023-03-19"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
runtime: shiny
---

## Load libraries
<!-- First, let's install and load the necessary packages - shiny, leaflet, and dplyr -->
```{r}
my_packages = c("plyr", "plotly", "ggplot2", "psych", "tidyr", "tidyverse","dplyr","lubridate","readr","caret","caTools","glmnet","shiny","shinyvalidate","shinyWidgets","shinythemes","tippy","shinyjs","leaflet","mapview")
#install.packages(my_packages)
lapply(my_packages, require, character.only = T)
```   

### Import the file *AB_NYC_2019.csv*
<!-- Next, we need to read the AB_NYC_2019.csv dataset into our Shiny app. For this, we can use the read.csv function. -->
```{r}
airbnb_unclean = read.csv("~/R_Projects/Data/AB_NYC_2019.csv")
print('Uncleaned Dataframe Info:')
cat("Number of Rows:", nrow(airbnb_unclean), "\n") # Printing string and variable row count on the same line
cat("Number of Columns:", ncol(airbnb_unclean), "\n") 
cat("Blank cells count:", sum(!complete.cases(airbnb_unclean))) # Displaying Blank Cells Count from the original data frame
headTail(airbnb_unclean, top = 3, bottom = 3, ellipsis = F)
```

### Dataframe Cleanup
```{r}
airbnb_sort= airbnb_unclean[order(-airbnb_unclean$price),] # Sorted dataframe with descending price

print('Removed Blank cells:')
airbnb_modified = na.omit(airbnb_sort)

print('Resized dataframe:')
airbnb = airbnb_modified[26892:22003,] # Remove the top 55 % and bottom 45%
cat("Number of Rows:", nrow(airbnb), "\n")
cat("Number of Columns:", ncol(airbnb), "\n")
cat("Blank cells count in cleaned dataframe:", sum(!complete.cases(airbnb)), "\n") # Displaying blank cells count after removal
headTail(airbnb, top = 3, bottom = 3, ellipsis = F) # Cleaned Dataframe
```

```{r}
# Load data.table package
library(data.table)
```

```{r}
# Write to CSV file
fwrite(airbnb, "~/R_Projects/Data/airbnb.csv")
```

## Now, we can create our Shiny app UI and server functions. In the UI function, we will create a leafletOutput object to display the choropleth map, and sliders to allow the user to filter the data based on the price and minimum_nights columns. -->
```{r}
ui <- fluidPage(
  titlePanel("Airbnb Listings in New York City"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("price_slider", "Filter by Price", min = 0, max = 1000, value = c(0, 500)),
      sliderInput("min_nights_slider", "Filter by Minimum Nights", min = 1, max = 30, value = c(1, 7))
    ),
    mainPanel(
      leafletOutput("map")
    )
  )
)

```


## In the server function, we will filter the airbnb based on the user's selected price and minimum nights values, and create a leaflet object with a choropleth layer to display the data on a map.
```{r}
server <- function(input, output) {
  
  filtered_data <- reactive({
    airbnb %>%
      filter(price >= input$price_slider[1] & price <= input$price_slider[2] & minimum_nights >= input$min_nights_slider[1] & minimum_nights <= input$min_nights_slider[2])
  })
  
  output$map <- renderLeaflet({
    leaflet(filtered_data()) %>%
      addTiles() %>%
      setView(lng = -73.987056, lat = 40.727248, zoom = 11) %>%
      addProviderTiles("CartoDB.Positron") %>%
      addMarkers(lng = ~longitude, lat = ~latitude, popup = ~name)
  })
}

```

## Finally, we can run the Shiny app using the shinyApp function. -->
<!-- This will open a Shiny app in web browser, where we can use the sliders to filter the Airbnb data based on price and minimum nights, and view the filtered data on a choropleth map of New York City. -->
```{r}
shinyApp(ui, server)
```


