---
title: 'ALY6070: Final Project'
subtitle: 'Airbnb New York Analysis (2011 - 2019)'
author: "Abhilash Dikshit"
date: "2023-04-03"
bibliography: /Users/abidikshit/R_Language/bibliography.bib
biblio-style: "apalike"
link-citations: true
output:
  
  pdf_document: default
  html_document:
    df_print: paged
---

# R Code Approach
### Importing packages: *FSA, FSAdata, magrittr, dplyr, tidyr, plyr and tidyverse*
```{r}
# install.packages("plyr")   # Import Libraries
# install.packages("dplyr")
# install.packages("tidyr")
# install.packages("tidyverse")
# install.packages("psych")
# install.packages("ggpubr")
# install.packages("ggplot2")
# install.packages("scales")

library(plyr) 
library(dplyr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(psych)
library(moments)
library(scales)
```

### Import the file *AB_NYC_2019.csv* and show basic info
```{r}
airbnb = read.csv("~/R_Projects/Data/airbnb.csv")
cat("Number of Rows:", nrow(airbnb), "\n") # Printing string and variable row count on the same line
cat("Number of Columns:", ncol(airbnb), "\n") 
cat("Blank cells count:", sum(!complete.cases(airbnb))) # Displaying Blank Cells Count from the original data frame
headTail(airbnb, top = 3, bottom = 3, ellipsis = F)
```

## Bar chart showing the distribution Of Room Types In New York City 
```{r}
counts <- airbnb %>% 
  count(room_type)

ggplot(data = counts) + 
  geom_bar(mapping = aes(x = room_type, y = n, fill = n), 
           stat = "identity", position = "dodge") + 
  scale_fill_gradient(low = "#005E7D", high = "#FFD662", guide = "legend") + 
  geom_text(aes(x = room_type, y = n, label = n), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 4.5) +
  labs(title = "Distribution Of Room Types In New York City", 
       x = "Room Type", y = "Count")
```

## Stacked bar chart showing the distribution of room types by neighbourhood group:
```{r}
ggplot(data = airbnb) + 
  geom_bar(mapping = aes(x = neighbourhood_group, fill = room_type), 
           position = "stack") +
  stat_count(mapping = aes(x = neighbourhood_group, label = stat(count)), 
             geom = "text", position = position_stack(vjust = 0.5), size = 3) +
  scale_fill_manual(values = c("#FFD700", "#56B4E9", "#7FFFD4")) +
  labs(title = "Room Type Distribution by Neighbourhood Group", 
       x = "Neighbourhood Group", y = "Count") +
  guides(fill = guide_legend(title = "Room Type"))
```


## Scatterplot showing the relationship between price and minimum nights:
```{r}
ggplot(data = airbnb) + 
  geom_point(mapping = aes(x = minimum_nights, y = price, color = neighbourhood_group), alpha = 0.5) + 
  labs(title = "Price vs. Minimum Nights", x = "Minimum Nights", y = "Price")
```

## Boxplot showing the distribution of prices by room type:
```{r}
ggplot(data = airbnb) + 
  geom_boxplot(mapping = aes(x = room_type, y = price, fill = room_type)) +
  scale_fill_manual(values = c("blue", "green", "red")) +
  labs(title = "Price Distribution by Room Type", x = "Room Type", y = "Price")
```

## Average Price For Each Room Type
```{r}
print("Clean Data Average Price For Each Room Type")
airbnb_room_mean = aggregate(price ~ room_type, airbnb, mean)
airbnb_room_mean

airbnb_pie = ggplot(airbnb_room_mean, aes(x= room_type, y= price, fill=room_type)) + geom_bar(stat="identity", width=1) +
    coord_polar("y", start=0)+scale_fill_brewer(palette="Dark2")+labs(title = "Average Price For Each Room Type", subtitle = "Cleaned Data", fill = "") + geom_text(aes(label = round(price, digits = 1)))
airbnb_pie
```

## Density plot showing the distribution of prices:
```{r}

ggplot(data = airbnb, aes(x = price, fill = room_type)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73")) +
  labs(title = "Price Distribution for all Airbnbs in New York", x = "Price", y = "Density", fill = "Room Type")

ggplot(data = airbnb) + 
  geom_density(mapping = aes(x = price), fill = "#56B4E9", alpha = 0.5) +
  labs(title = "Price Distribution", x = "Price", y = "Density")

ggplot(data = subset(airbnb, room_type == "Private room" & neighbourhood_group == "Manhattan")) + 
  geom_density(mapping = aes(x = price), fill = "#56B4E9", alpha = 0.5) +
  labs(title = "Price Distribution for Private Rooms in Manhattan", x = "Price", y = "Density")

```

## Mapview: Number Of Reviews Density Using Latitude and Longitude For Specific Room Types As Per Price
```{r}
mapview(airbnb_unclean, xcol = "longitude", ycol = "latitude", crs = 4326, grid = FALSE) # coordinate reference system (crs)
```

```{r}
qmplot(longitude, latitude, data = airbnb, geom = "point", color = price, main = "Cleaned Data: Density of Latitude and Longitude") + facet_wrap(~ room_type)
```

# Shiny R Approach

## Scatter Plot
```{r}
library(shiny)
library(ggplot2)
library(plotly)
library(dplyr)

ui <- fluidPage(
  titlePanel("Airbnb Listings in New York City, USA (2011-2019)"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("n", "Number of observations to view:", 
                  min = 1, max = nrow(airbnb), value = 50)
    ),
    mainPanel(
      plotlyOutput("plot")
    )
  )
)

server <- function(input, output) {
  output$plot <- renderPlotly({
    gg <- ggplot(airbnb[1:input$n,], aes(x = minimum_nights, y = price, color = neighbourhood_group, label = name)) +
      geom_point(alpha = 0.5) +
      ggtitle("Scatterplot of Price and Minimum Nights") +
      labs(x = "Minimum Nights", y = "Price") +
      theme_bw()
    ggplotly(gg, tooltip = c("name"))
  })
}

shinyApp(ui, server)

```

## Boxplot
```{r}
library(shiny)
library(plotly)
library(dplyr)

ui <- fluidPage(
  titlePanel("Airbnb Listings in New York City, USA (2011-2019)"),
  sidebarLayout(
    sidebarPanel(
      selectInput("room_type", "Choose a Room Type:",
                  choices = c("All", unique(airbnb$room_type))), # added "All" option as default
      sliderInput("price", "Filter by Price:", 
                  min = 0, max = max(airbnb$price), value = c(0, max(airbnb$price))),
      sliderInput("nights", "Filter by Minimum Nights:", 
                  min = 1, max = 30, value = c(1, max(airbnb$minimum_nights)))
    ),
    mainPanel(
      plotlyOutput("boxplot")
    )
  )
)

server <- function(input, output) {
  
  filtered_airbnb <- reactive({
    if(input$room_type == "All") {
      airbnb %>%
        filter(price >= input$price[1] & price <= input$price[2] &
                 minimum_nights >= input$nights[1] & minimum_nights <= input$nights[2])
    } else {
      airbnb %>%
        filter(room_type == input$room_type & 
                 price >= input$price[1] & price <= input$price[2] &
                 minimum_nights >= input$nights[1] & minimum_nights <= input$nights[2])
    }
  })
  
  output$boxplot <- renderPlotly({
    p <- ggplot(filtered_airbnb(), aes(x = room_type, y = price, fill = room_type)) +
      geom_boxplot(alpha = 0.5) +
      ggtitle(paste0("Boxplot of ", input$room_type, " Room Type and Price")) +
      xlab("Room Type") + ylab("Price") +
      theme_minimal()
    ggplotly(p, tooltip = c("y", "lower", "middle", "upper", "outlier"), 
              width = 800, height = 500) %>%
      layout(showlegend = FALSE)
  })
  
}

shinyApp(ui, server)
```


### Stacked bar chart
```{r}
library(shiny)
library(ggplot2)
library(plotly)
library(dplyr)
library(magrittr)

ui <- fluidPage(
  titlePanel("Airbnb Listings in New York City, USA (2011-2019)"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("n", "Number of observations to view:", 
                  min = 1, max = nrow(airbnb), value = 50),
      checkboxGroupInput("room_type", "Select Room Types to Display:",
                         choices = unique(airbnb$room_type),
                         selected = unique(airbnb$room_type)),
      selectInput("neighbourhood_group", "Select a Neighbourhood Group:",
                  choices = c("All", unique(airbnb$neighbourhood_group)))
    ),
    mainPanel(
      plotlyOutput("plot")
    )
  )
)

server <- function(input, output) {
  
  filtered_airbnb <- reactive({
    if(input$neighbourhood_group == "All"){
      airbnb %>%
        filter(room_type %in% input$room_type) %>%
        slice(1:input$n)
    } else {
      airbnb %>%
        filter(neighbourhood_group == input$neighbourhood_group &
                 room_type %in% input$room_type) %>%
        slice(1:input$n)
    }
  })
  
  output$plot <- renderPlotly({
    filtered_airbnb() %>%
      ggplot(aes(x = neighbourhood, fill = room_type)) +
      geom_bar() +
      labs(title = "Stacked Bar chart of Neighbourhood and Room Type") +
      theme(plot.title = element_text(size = 20)) +
      scale_fill_brewer(palette = "Set1") +
      theme(legend.title = element_blank(),
            legend.position = "bottom",
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -> p
      
    ggplotly(airbnb = p %+% airbnb.frame(neighbourhood = filtered_airbnb()$neighbourhood), 
             tooltip = c("room_type", "count")) %>%
      layout(hovermode = "closest",
             hoverlabel = list(bgcolor = "white", 
                               font = list(family = "Roboto", size = 14),
                               align = "left"),
             xaxis = list(title = "Neighbourhood", tickfont = list(size = 12)),
             yaxis = list(title = "Count", tickfont = list(size = 12)))
  })
}

shinyApp(ui, server)
```

## Histogram plot
```{r}
library(shiny)
library(plotly)

# Define colors for each room_type
colors <- c("#7fc97f", "#beaed4", "#fdc086", "#ffff99")

ui <- fluidPage(
  titlePanel("Price Distribution of Airbnb Listings in New York City"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("price_range", "Price Range:", 
                  min = 0, max = 1000, value = c(0, 500)),
      br(),
      checkboxGroupInput("room_type", "Select Room Types to Display:",
                         choices = unique(airbnb$room_type),
                         selected = unique(airbnb$room_type)),
      br(),
      selectInput("neighbourhood_group", "Select a Neighbourhood Group:",
                  choices = c("All", sort(unique(airbnb$neighbourhood_group)))),
      br(),
      selectInput("minimum_nights", "Minimum Nights:",
            choices = c("All", 1:30),
            selected = "All"),
      br(),
      selectInput("availability_365", "Availability 365:",
                  choices = c("All", sort(unique(airbnb$availability_365))))
    ),
    mainPanel(
      plotlyOutput("plot")
    )
  )
)


server <- function(input, output) {
  
  filtered_airbnb <- reactive({
    # Map each room_type to a color
    colors_map <- setNames(colors, unique(airbnb$room_type))
    
    airbnb %>%
      filter(price >= input$price_range[1] & price <= input$price_range[2],
             room_type %in% input$room_type,
             if(input$neighbourhood_group == "All") TRUE else neighbourhood_group == input$neighbourhood_group,
             if(input$minimum_nights == "All") TRUE else minimum_nights == as.numeric(input$minimum_nights),
             if(input$availability_365 == "All") TRUE else availability_365 == as.numeric(input$availability_365)
             ) %>%
      # Add a new column "color" based on room_type
      mutate(color = colors_map[room_type])
  })
  
  output$plot <- renderPlotly({
    filtered_airbnb() %>%
      plot_ly(x = ~price, type = "histogram",
              marker = list(color = ~color, 
                            line = list(color = "white", width = 0.5)),
              hovertemplate = paste(
                "<b>Price:</b> %{x:.0f}<br>",
                "<b>Count:</b> %{y:.0f}<br>",
                "<b>Room Type:</b> %{marker.color}"
              ),
              showlegend = FALSE
              ) %>%
      layout(xaxis = list(title = "Price"),
             yaxis = list(title = "Count"),
             hovermode = "closest",
             hoverlabel = list(bgcolor = "white", 
                               font = list(family = "Roboto", size = 14),
                               align = "left"),
             colorway = colors,
             margin = list(l = 40, r = 40, t = 40, b = 40))
  })
}

shinyApp(ui, server)

```

## Mapview
```{r}
library(leaflet)

ui <- fluidPage(
  titlePanel("Airbnb Listings in New York City"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("price_slider", "Filter by Price", min = 0, max = 1000, value = c(0, 500)),
      sliderInput("min_nights_slider", "Filter by Minimum Nights", min = 1, max = 30, value = c(1, 7))
    ),
    mainPanel(
      leafletOutput("map")
    )
  )
)

server <- function(input, output) {
  
  filtered_airbnb <- reactive({
    airbnb %>%
      filter(price >= input$price_slider[1] & price <= input$price_slider[2] & minimum_nights >= input$min_nights_slider[1] & minimum_nights <= input$min_nights_slider[2])
  })
  
  output$map <- renderLeaflet({
    pal <- colorFactor(palette = "Dark2", domain = filtered_airbnb()$room_type)
    leaflet(filtered_airbnb()) %>%
      addTiles() %>%
      setView(lng = -73.987056, lat = 40.727248, zoom = 11) %>%
      addProviderTiles("CartoDB.Positron") %>%
      addCircleMarkers(lng = ~longitude, lat = ~latitude, radius = 5, 
                       color = ~pal(room_type), fillOpacity = 0.8, stroke = FALSE,
                       popup = ~paste("<b>Name:</b> ", name, "<br>",
                                      "<b>Room Type:</b> ", room_type, "<br>",
                                      "<b>Price:</b> $", price, "/night"))
  })
}

shinyApp(ui, server)

```




